\chapter{Commentary and Conclusions}

The \textbf{\acrshort{SVP64} \acrshort{POWER} ISA Vector Optimisation for Search}
project was conceived by RED Semiconductor, VectorCamp and VanTosh to build on
a body of previously published open-source development work that had proposed
extensions to the OpenPOWER \acrfull{ISA} that would make it more powerful and
efficient for algorithmic computation such as that utilised in internet search
applications. It is envisioned that optimisation of search will bring benefits
in speed and accuracy of searches, and reduce the power consumption required to
perform them.
Furthermore, efficiency of algorithmic computation is expected to deliver
security benefits in terms of personal privacy and data security
across the internet.

The collaborators have a broad spectrum of relevant technical knowledge, as
well as being businesses which take computing concepts into real world
applications:
\begin{itemize}
  \item RED Semiconductor is a developer of computing IP solutions that embrace
  microprocessor hardware architecture features and the command-and-control
  software that invokes them. RED researches incumbent architectures and
  creates new additional features to overcome their limitations in regard to
  processing performance for next generation demands.
  \item VectorCamp specialises in optimising software utilising \acrshort{SIMD}
  and other optimisation techniques, to deliver performance improvements on
  hardware platforms including Arm, Intel and \acrshort{POWER}. VectorCamp has
  developed a high performance pattern matching software tool called Vectorscan.
  \item VanTosh is a provider of IT Services, and has full-stack expertise
  internet and search ecosystems. VanTosh offers software development and
  operations services designed to optimised the effectiveness of IT use.
\end{itemize}

The combination of RED’s expertise in hardware, VectorCamp’s in software and
VanTosh’s in IT ecosystems makes the consortium ideally placed for research and
development into new optimised solutions for internet search, and the three
organisations have previously been involved in projects concerning
\acrshort{SVP64} and \acrshort{POWER} ISA.

Key objectives of the
\textbf{\acrshort{SVP64} \acrshort{POWER} ISA Vector Optimisation for Search}
project, undertaken by the consortium of RED, VectorCamp, and VanTosh
has had the following objectives:

\begin{itemize}
  \item Explore the practical capabilities of \acrshort{SVP64} in its current
        iteration for internet search applications
  \item Research and plan necessary work required to “De-SIMDfy” the Vectorscan
        code base in order to make it scalar-portable
  \item Write experimental string/memory routines using \acrshort{SVP64}
        assembler
  \item Integrate the above routines within a standard C library test framework
  \item Document and report \acrshort{SVP64} and its simulator for future
        developers and users (and report back on issues to developers of
        \acrshort{SV}, if any occurred)
  \item Research common \acrshort{SIMD} routines used in Vectorscan, and work
        on draft implementations in \acrshort{SVP64}
        (enables Vectorscan support for scalar architectures)
  \item Report on how \acrshort{SVP64} can be applied to optimise search for
        the benefit of future users and projects
\end{itemize}

Each of the objectives and tasks was addressed and completed during the
project, and this report gives detailed descriptions of those tasks and the
software code created in order to learn about and demonstrate \acrshort{SVP64}’s
usefulness in optimising search algorithms.
In the process of completing the project a number of challenges were
encountered and largely overcome in relation to the current stage of
development of \acrshort{SVP64} and its readiness to be applied to microprocessor
hardware architectures.
The consortium has concluded that in concept \acrshort{SVP64} permits software developers
to optimise a wide array of algorithms, including for applications other
than search, but in order for realistic hardware implementation
further development is required.

\acrfull{SV} (\acrshort{SVP64} being a particular variant for \acrshort{POWER}
ISA) is a powerful extension on top of scalar instructions, if the software
developer understands the set of features it contains, and how to use it.

To demonstrate search pattern algorithms using \acrshort{SVP64}, it was decided
to integrate code written in this extension into the \texttt{glibc}
test framework.
Although the present state of SVP64's development and maturity has enabled RED,
VectorCamp and VanTosh to undertake specific algorithm code developments for
evaluation and demonstration purposes, this project has taught us that
\acrshort{SV} still requires further development to make is a practicable
extension to the POWER ISA. Moreover, in order to realise the benefits of SVP64
in a real-world environment, its ISA-based features require hardware
developments to the microprocessor architecture.

\section{A Hardware Perspective by Red Semiconductor}

Red Semiconductor has concluded from this project that SVP64 has reached a
stage of its development where hardware realisation must be carefully
considered. Research undertaken during this project has revealed that whereas
certain SVP64 features can be implemented in a straightforward manner, there
are many that require investigation by chip architects with microprocessor
design expertise in order to specify and implement their functionality.

Although for this project it has been beneficial to abstract away hardware
specifics from the programmer, it is still necessary to consider how many of
\acrshort{SV}'s features can actually be implemented in reality,
balancing the overall performance benefits with practical issues like the
complexity of the hardware needed, and its cost in silicon area and power
consumption when compared with other ISAs and Microprocessor chips.

Many advanced features of SVP64, such as the REMAP system, fail-first,
predication, element-width overrides etc., are currently difficult to use
due to the fact that the reference simulator for the SVP64 specification is
still in the experimental phase of its development, and during this project the
code writers discovered that its functionality, and doesn't always align with
the available documentation.

This issue will be fed back to the individuals who developed the simulator in
order to guide their further work. Nevertheless, a lot of time was spent
during Milestones 2 and 3 simply getting around the issues of the simulator,
and the available assembler notation, and this limited the scope of function
complexity and optimisation, but we believe it will be valuable 'user' input
to its developers, enabling them to create a better functioning tool
for future use.

\section{A Software Perspective by VectorCamp}
\label{sec:vecotorscansoftwareperspective}

Existing code base for many projects utilising vector extensions are on the
order of 100k lines of code or more.

Thus it is unreasonable to refactor the entire codebase using a new vector
extension, regardless of how performant it may be.

Instead, it is better to gradually upgrade parts of the code (just as is
happening today with the Rust programming language), by substituting specific
vector extension (AVX, NEON, etc.) functions with optimised SVP64 code instead.

This is why during Milestone 3, we went through the exercise of researching
common SIMD operations, and implementing them in SVP64. By having
implementations of standard operations (such as shuffle, movemask, etc.)
gradual refactoring can be accomplished in the future.

However, due to the lack of specific byte manipulation instructions, and
difficulty in configuration setup of the REMAP system, these implementations
turned to be either much bigger than desired, or only partially implemented.


\section{An ISA \& Distribution Perspective by VanTosh}

After understanding SVP64 much better, it is apparent that the extension has
the possiblity to introduce new concepts to the \acrshort{POWER} \acrshort{ISA},
however the disadvantages outway the advantages, while we have been able to
reduce the number of instructions, the legiability of code and the maintanability
becomes much more complex.
After careful consideration there are simpler ways to achieve several optimizations
SVP64 highlights and that can be added to the \acrshort{POWER} \acrshort{ISA},
such as Vector Element Move, Mask or Shift.

From a higher software ecosystem, without operating system support or library support
SVP64, as suggested in \label{sec:vecotorscansoftwareperspective} by VectorCamp,
would be cumbersome to convert and would be useless.
In our previous efforts to get Linux Distribution support to SVP64, our effort to
get a minimal \acrshort{POWER} architecture ``ppc64sffs" running, we were able
to get the base system supported for \href{https://www.powerel.org}{PowerEL},
\href{https://www.debian.org}{Debian}, and \href{https://gentoo.org}{Gentoo}.
However we never were able to verify the \acrshort{POWER} \acrfull{SFFS}
The benefit of adding SVP64 never got proven in detail and while this project
has been able to prove a better cycle usage to reduce the number of instruction
cycles, it does not yet prove the increase of speed that would be inferred.

\section{Future Work}

This project has been the first attempt by attempt by multiple organisations
to utilise the capabilities embodied in the SVP64 concept in real-world
applications, and much has been learnt about the innovative features as well
as the challenges and limitations of applying SVP64 to existing
coding methodology and applications.

Based on the experience of the consortium, Red has learned enough to move from
the purely theoretical towards a real functional HW specification
and implementation.
In practice, this has caused us to have a fundamental rethink of the concepts
realised in SVP64, and has led to us going back to the original theoretical
ideas that inspired its development in order to develop a practicable hardware
solution, which will differ from \acrshort{SV}, in order to deliver
advanced functionality that is required by developers of systems for next
generation internet search and infrastructure projects where key driving force
is likely to include the use of AI for optimised search, and cryptography to
maintain users' privacy and security of data.

The NGI Search collaborative project has proved to be a successful knowledge
platform for the contributors, and has started to make connections between the
deep research history behind the SVP64 concept and its potential use in
real-world applications. The teams have experienced the challenges associated
with moving the excellent theoretical concepts previously demonstrated in
isolated code snippets into the realm of practical usage within large code
bases comprising many complex algorithms. The project has reinforced the
inspirations behind SVP64 and it raison d’être, while exposing the challenges
associated with its applicability to the POWER ISA, and its ease of use when
integrating with traditional coding methodologies. The collaborators have
already considered future development work, including the alignment of future
ISA extensions with pre-existing performance-boosting methodologies like SIMD,
and how a hardware architecture can then be implemented to optimise for the
performance benefits that can be obtained from the modified ISA.

It is highly likely that the members of this project will collaborate on future
developments and seek grant funding where appropriate. The teams also hope that
this published work will inspire others individuals and organisations to join
the consortium or work independently on future vector extensions.

We would like to thank the Libre-SOC project.
